Exploration of Red Wine Quality by Jeremy Yenko
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
```

```{r echo=TRUE, Load_the_Data}
rw <- read.csv('wineQualityReds.csv')

```

This analysis will focus on the many chemical properties that have the potential
to influence the quality of red wines.    

# Preliminary exploration

### Variables in our dataset

```{r echo=TRUE, Univariate_Plots}
str(rw)

```

Above, we can see a number of chemical factors present that can influence the
quality of red wine.  

At the bottom of this list lies our independent variable: quality. 

Quality is measured on a scale of 0-10 with ratings being provided by a minimum 
of three wine experts for each wine.

### Overview of wine quality

```{r echo=TRUE, message=FALSE, warning=FALSE, First_Plot}
qplot(data = rw, x = quality) +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 1))
```

As seen above, the quality ratings of the wines in our study appear to have
a normal distribution.

## Univariate Analysis

### Alcohol content and wine quality

> One of the first aspects of red wine we will want to examine is alcohol content.
Below is a summary of the alcohol content of the wines in our dataset:

```{r echo=TRUE, message=FALSE, warning=FALSE, Alcohol_Content_Summary}
summary(rw$alcohol)
```

As seen above, our wines have a mean alcohol content of 10.42.  Since the mean
is close to the median, this suggests we do not have many outliers and that 
the mean is generally a reliable measure for the alcohol content variable.

```{r echo=TRUE, message=FALSE, warning=FALSE, Basic_Info}
rw_hq <- subset(rw, quality >= 8)
qplot(data = rw_hq, x=alcohol, bins=50, main="Alcohol in wines rated 8 and above") +
  xlim(8,15)
rw_lq <- subset(rw, quality <= 4)
qplot(data=rw_lq, x=alcohol, bins=50, main="Alcohol in wines rated 4 and below") +
  xlim(8,15)
qplot(data = rw, x=alcohol, bins=50, main="Alcohol in all wines") +
  xlim(8,15)
```

These two plots show that the higher quality wines tend to have a greater 
alcohol content than the lower quality wines.  For all wines, the distribution
appears positively skewed.

Now we will examine the correlation between alcohol and wine quality:

```{r echo=TRUE, message=FALSE, warning=FALSE}
cor.test(rw$alcohol, rw$quality)
```

We can see that there is a positive correlation between alcohol content and wine
quality.  Although this correlation is not notably strong, the low p-value
suggests it is a reliable measure.

