Exploration of Red Wine Quality by Jeremy Yenko
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
```

```{r echo=TRUE, Load_the_Data}
rw <- read.csv('wineQualityReds.csv')

```

This analysis will focus on the many chemical properties that have the potential
to influence the quality of red wines.    

# Preliminary exploration

### Variables in our dataset

```{r echo=TRUE, Univariate_Plots}
str(rw)

```

Above, we can see a number of chemical factors present that can influence the
quality of red wine.  

At the bottom of this list lies our independent variable: quality. 

Quality is measured on a scale of 0-10 with ratings being provided by a minimum 
of three wine experts for each wine.

### Overview of wine quality

```{r echo=TRUE, message=FALSE, warning=FALSE, First_Plot}
qplot(data = rw, x = quality) +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 1))
```

As seen above, the quality ratings of the wines in our study appear to have
a normal distribution.

## Univariate Analysis

### Alcohol content and wine quality

> One of the first aspects of red wine we will want to examine is alcohol content.
Below is a summary of the alcohol content of the wines in our dataset:

```{r echo=TRUE, message=FALSE, warning=FALSE, Alcohol_Content_Summary}
summary(rw$alcohol)
```

As seen above, our wines have a mean alcohol content of 10.42.  Since the mean
is close to the median, this suggests we do not have many outliers and that 
the mean is generally a reliable measure for the alcohol content variable.

```{r echo=TRUE, message=FALSE, warning=FALSE, Basic_Info}
rw_hq <- subset(rw, quality >= 8)
qplot(data = rw_hq, x=alcohol, bins=50, main="Alcohol in wines rated 8 and above") +
  xlim(8,15)
rw_lq <- subset(rw, quality <= 4)
qplot(data=rw_lq, x=alcohol, bins=50, main="Alcohol in wines rated 4 and below") +
  xlim(8,15)
qplot(data = rw, x=alcohol, bins=50, main="Alcohol in all wines") +
  xlim(8,15)
```

These two plots show that the higher quality wines tend to have a greater 
alcohol content than the lower quality wines.  For all wines, the distribution
appears positively skewed.

Now we will examine the correlation between alcohol and wine quality:

```{r echo=TRUE, message=FALSE, warning=FALSE}
cor.test(rw$alcohol, rw$quality)
```

We can see that there is a positive correlation between alcohol content and wine
quality.  Although this correlation is not notably strong, the low p-value
suggests it is a reliable measure.

### pH and wine quality

> Another factor that can influence wine quality is pH.  Since wine is generally
acidic, we will investigate whether or not the level of acidity influences wine
quality.

First, we will look at the distribution of pH across all of the wines in our 
dataset.

```{r echo=TRUE, message=FALSE, warning=FALSE}
qplot(data = rw, x=pH, bins=50, main="Distribution of pH across all wines")
```

The pH values of our wines appear to follow a normal distribution.

```{r echo=TRUE, message=FALSE, warning=FALSE}
summary(rw$pH)
```

All of our wines have a pH well below 7.  Our most acidic wine has a pH value of
2.740 and our least acidic wine has a pH value of 4.010.  The mean and median 
are close, indicating the mean is generally a reliable measure for the pH of
our wines.

We will now investigate the relationship between pH and quality.

```{r echo=TRUE, message=FALSE, warning=FALSE}
rw_pH_hq <- subset(rw, quality>=8)
rw_pH_lq <- subset(rw, quality<=4)

qplot(data=rw_pH_hq, x=pH, bins=50) +
  xlim(2.7,4.1) +
  ylim(0,10)
qplot(data=rw_pH_lq, x=pH, bins=50) +
  xlim(2.7,4.1) +
  ylim(0,10)
```